\plaintitle

### 2025-05-21 docs/style/article-body.tex
for incoming pull

\textbf{Iteration 6 - Data type exploration:} Following the removal of vectors, this iteration explored alternative static data types to determine the most efficient structure for storing input and seed data.

\textbf{Iteration 7 - Re-implementation of bitshifting:} In this final iteration, bit shifting techniques were reintroduced into the now simplified implementation. With high-level abstractions removed and memory structures streamlined, bit-level operations such as shifting and masking were applied.

@@

\begin{tabular}{|c|c|c|c|c|c|c|} \hline \textbf{Bit size} &
\multicolumn{3}{c|}{\textbf{Teensy ($\mu s$)}} &
\multicolumn{3}{c|}{\textbf{Pico ($\mu s$)}} \\ \hline & Min & Max & Avg & Min &
Max & Avg \\ \hline 64 & 13 & 14 & 13.1564 & 103 & 117 & 106.3914 \\ 512 & 788 &
791 & 788.3139 & 5296 & 5311 & 5302.4979 \\ 1024 & 3124 & 3130 & 3124.0580 &
21108 & 21119 & 21111.2163 \\ \hline \end{tabular} \caption{Minimum, maximum and
average processing time for Teensy and Pico.}
\label{tab:teensy_pico_min_max_avg}

Table 1 presents the execution speeds of each iteration as listed in sections 5.3 and 5.4, alongside the execution speed in Mbit/s as calculated by utilizing the average value in (3). My thought then is that we here talk about the results and what they demonstrate, before moving on to draw conclusions and attempt to answer our research questions in section 7.

>>>>

\textbf{Iteration 6 - Data type exploration:} Following the removal of vectors, this iteration explored alternative static data types such as: \texttt{array}, \texttt{unordered\_map} and \texttt{bitset} to determine the most efficient structure for storing input and seed data.

@@

\subsubsection{6.1 Phase one}\label{phase-one-1}

\textbf{Iteration 1 - Data Structures:}

\begin{tabular}{|c|c|c|} \hline \textbf{Bit size} &
\multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}} &
\multicolumn{1}{c|}{\textbf{Pico ($\mu s$)}} \\ \hline 64 & 13.1564 & 106.3914
\\ 512 & 788.3139 & 5302.4979 \\ 1024 & 3124.0580 & 21111.2163 \\ \hline
\caption{A brief overview of the testscript architecture.} \label{tab:iter1}

\end{tabular}

Table \ref{tab:iter1} presents the average execution time of iteration 1 on both Teensy 4.1 and Raspberry Pico Pi 2 across thre input sizes. The Teensy consistently outperforms the Pico, with the gap widening as the bit size increases.

\textbf{Iteration 2 - Bitshifting:}

\begin{tabular}{|c|c|} \hline \textbf{Bit size} &
\multicolumn{1}{c|}{\textbf{Avg ($\mu s$)}} \\ \hline 64 & 16.4689 \\ 512 &
1006.6255 \\ 1024 & 3996.5972 \\ \hline \end{tabular}

\label{tab:iter2}

Table \ref{tab:iter2} shows the execution times for iteration 2 on the Teensy 4.1, which utilizes bitshifting instead of the original data structure approach. A clear increase in executiom time compared to iteration 1 is observed.

Additionally, the isolated effect of applying a single bitmask operation \texttt{\&\ 1} was evaluated. This resulted in a slight reduction in average execution time, from \texttt{13.1564\ µs} to \texttt{13.1000\ µs}, although this specific result is not included in the tables above.

\textbf{Iteration 3 - Batching and Hardware optimization:}

\begin{tabular}{|c|c|} \hline \textbf{Bit size} &
\multicolumn{1}{c|}{\textbf{Avg ($\mu s$)}} \\ \hline 64/64 & 43.0760 \\ 512/512
& 2663.8194 \\ 1024/1024 & 10513.1767 \\ \hline \end{tabular}

\label{tab:iter3}

Table \ref{tab:iter3} presents the average execution time of iteration 3 on the Teensy 4.1 for varying batching and bit sizes. Additionally, when performing the same 1024/1024-bit operation, the Raspberry Pi Pico 2 showed average execution time of \texttt{80806.006\ µs} indicating a substantially lower throughput compared to the Teensy.

\subsubsection{6.2 Phase two}\label{phase-two-1}

\textbf{Iteration 4 - Loop unrolling:}

\begin{tabular}{|c|c|c|} \hline \textbf{Bit size} &
\multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}} &
\multicolumn{1}{c|}{\textbf{Pico ($\mu s$)}} \\ \hline 64 & 9.7017 & 70.7402 \\
512 & 551.6358 & 3978.1928 \\ 1024 & 2195.1979 & 15830.8784 \\ \hline
\end{tabular}

\label{tab:iter4}

Table \ref{tab:iter4} presents the execution time of iteration 4, which utilizes a 4x unrolled bit-processing loop. Compared to erlier iterations, this approach yields a substantial performance improvments on both the Teensy and the Raspberry Pi Pico Mcu's.

In addition to the results presented in Table \ref{tab:iter4}, a further tests was conducted on the Teensy using a single-loop unrolled implementation for the 64-bit size, resulting in an average execution time of \texttt{6.6626\ µs}. An additional fully unrolled variant, where loops were entirely eliminated, produced a measured execution time of \texttt{0.0491\ µs}. However, this measurement was later determined to be invalid due to packaging error.

\textbf{Iteration 5 - Removal of vector usage:}

\begin{tabular}{|c|c|c|} \hline \textbf{Bit size} &
\multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}} &
\multicolumn{1}{c|}{\textbf{Pico ($\mu s$)}} \\ \hline 64 & 0.0501 & 0.2175 \\
\hline \end{tabular}

\label{tab:iter5}

Table \ref{tab:iter5} presents the results of iteration 5, in which vector structure were removed in favor of fixed-size integer types \texttt{uint32\_t} and \texttt{uint64\_t}. This iteration led to an extremely efficient implementation that approaches the physical execution limits of the Teensy 4.1.

\textbf{Iteration 6 - Data type exploration:}

\begin{tabular}{|c|c|} \hline \textbf{Data structure} &
\multicolumn{1}{c|}{\textbf{Avg ($\mu s$)}} \\ \hline array & 0.4284 \\
unordered_map & 31.5090 \\ bitset & 0.0474 \\ \hline \end{tabular}

\label{tab:iter6}

Table \ref{tab:iter6} presents the results of iteration 6, which was introduced to address the structural limitations encountered in iteration 5. Specifically, the fixed-width output constrains imposed by the removal of vector. In this iteration, alternative data structures were evaluated. Vector-based approach from iteration 1 serving as a baseline for comparison.
